---
title: "Dev_between_sensors"
author: "Flueckiger_Rubens"
date: "2025-04-01"
output: 
  html_document:
    toc: true
---

# 1 Deviation between Sensors

In this file differences between a reference sensor to the others are visualised Here I took Sensor_3 (3m above ground) as reference Sensor. The choice roots from measurement conventions for devices at this height. Additionally, Bias-correction was performed at the same height, whereas a cross-validation between all sensors where not performed.

*Note:* Due to the moist adiabatic lapse rate being at Ts = 0.0069K/m (for T = 273.2K or 0°C) and Ts = 0.0036K/m (for T = 303K or 29.8°C) [Pal, 2001, P.68], the decrease in Temperature is negligible (in our case: 3m \* 0.0036K/m = 0.0108K decrease) and high degree differences (might) only occur due to device inaccuracy/bias.

Arya, S. Pal (2001), Introduction to Micrometeorology.Vereinigtes Königreich,Elsevier Science,2001.

# 2 Import packages and file

```{r, warning=FALSE, message=FALSE}
#Packages
library(dplyr)
library(ggplot2)

#File
lap <- read.table("../data/Distanz_&_Hohe_GR.csv", sep = ";", header = T)
```

# 3 New dataset

In this step a new dataset *anomaly* gets created. The file consists of:

-   Messung: This is the number of measurment
-   Reference sensor: Sensor_1 as reference (all values are = 0)
-   Remaining sensors: sensor - reference sensor
-   Both reference sensor and remaining sensors for temperature as well as for relative air humidity
-   Distanz_Runsum: Distance from measurment start in meters

```{r, echo=FALSE}
#Creating data frame
Messung <- lap$Messung
anomaly <- data.frame(Messung)

#Adding Values for temperature
anomaly <- mutate(anomaly, Sensor1_T = lap$Sensor1_T - lap$Sensor4_T) %>% #Sensor_0.5
  mutate(anomaly, Sensor2_T = lap$Sensor2_T - lap$Sensor4_T) %>% #Sensor_1
  mutate(anomaly, Sensor3_T = lap$Sensor3_T - lap$Sensor4_T) %>% #Sensor_2
  mutate(anomaly, Sensor4_T = lap$Sensor4_T - lap$Sensor4_T) #Sensor_3

#Adding Values for relative air humidity
anomaly <- mutate(anomaly, Sensor1_LF = lap$Sensor1_LF - lap$Sensor4_LF) %>% #Sensor_0.5
  mutate(anomaly, Sensor2_LF = lap$Sensor2_LF - lap$Sensor4_LF) %>% #Sensor_1
  mutate(anomaly, Sensor3_LF = lap$Sensor3_LF - lap$Sensor4_LF) %>% #Sensor_2
  mutate(anomaly, Sensor4_LF = lap$Sensor4_LF - lap$Sensor4_LF) #Sensor_3

#Adding distance
anomaly <- mutate(anomaly, Distance = lap$Distanz_Runsum)
```

# 4 Plotting

color codes

```{r, echo=FALSE}
colors <- c("Sensor_0.5" = "red", "Sensor_1" = "orange", "Sensor_2" = "darkgreen", "Sensor_3" = "blue", "Instrumentenfehler" = "lightblue")
```

Temperature

```{r, echo=FALSE}
#Open jpeg
jpeg("../Images/Dev_between_sensors/temp.jpeg", height = 486, width = 780, res = 108)

ggplot(anomaly, aes(x = Distance)) + #x-Axis
  geom_ribbon(aes(ymin = -0.1, ymax = 0.1, fill = "Instrumentenfehler"), alpha = 1) + #Filling
  geom_line(aes(y = Sensor1_T, color = "Sensor_0.5")) + #Sensor_0.5
  geom_line(aes(y = Sensor2_T, color = "Sensor_1")) + #Sensor_1
  geom_line(aes(y = Sensor3_T, color = "Sensor_2")) + #Sensor_2
  geom_line(aes(y = Sensor4_T, color = "Sensor_3")) + #Sensor_3
  geom_line(aes(y = mean(Sensor1_T), color = "Sensor_0.5"), linetype = "dashed") + #Mean Sensor_0.5
  geom_line(aes(y = mean(Sensor2_T), color = "Sensor_1"), linetype = "dashed") + #Mean Sensor_1
  geom_line(aes(y = mean(Sensor3_T), color = "Sensor_2"), linetype = "dashed") + #Mean Sensor_2
  geom_line(aes(y = mean(Sensor4_T), color = "Sensor_3"), linetype = "dashed") + #Mean Sensor_3
  labs(x = "Distanz [m]", y = "Temperaturdifferenz [°C]", #Axis Label
       color = "Legende", fill = "") + #Legend
  scale_color_manual(values = colors) + #Coloring legend
  scale_fill_manual(values = colors) +
  ggtitle("Abweichung zum Sensor_3") + #Adding title
  theme_bw() + #Changing plot type
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Close and save jpeg
dev.off()
```

Relative air humidity

```{r, echo=FALSE}
#Open jpeg
jpeg("../Images/Dev_between_sensors/humidity.jpeg", height = 486, width = 780, res = 108)

ggplot(anomaly, aes(x = Distance)) + #x-Axis
  geom_ribbon(aes(ymin = -1.5, ymax = 1.5, fill = "Instrumentenfehler"), alpha = 1) + #Filling
  geom_line(aes(y = Sensor1_LF, color = "Sensor_0.5")) + #Sensor_0.5
  geom_line(aes(y = Sensor2_LF, color = "Sensor_1")) + #Sensor_1
  geom_line(aes(y = Sensor3_LF, color = "Sensor_2")) + #Sensor_2
  geom_line(aes(y = Sensor4_LF, color = "Sensor_3")) + #Sensor_3
  geom_line(aes(y = mean(Sensor1_LF), color = "Sensor_0.5"), linetype = "dashed") + #Mean Sensor_0.5
  geom_line(aes(y = mean(Sensor2_LF), color = "Sensor_1"), linetype = "dashed") + #Mean Sensor_1
  geom_line(aes(y = mean(Sensor3_LF), color = "Sensor_2"), linetype = "dashed") + #Mean Sensor_2
  geom_line(aes(y = mean(Sensor4_LF), color = "Sensor_3"), linetype = "dashed") + #Mean Sensor_3
  labs(x = "Distanz [m]", y = "Luftfeuchtigkeitsdifferenz [%]", #Axis label
    color = "Legende", fill = "") + #Legend
  scale_color_manual(values = colors) + #Coloring legend
  scale_fill_manual(values = colors) +
  ggtitle("Abweichung zum Sensor_3") + #Adding title
  theme_bw() + #Changing plot type
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Close and save jpeg
dev.off()
```

# 4 Statistics

```{r, echo=FALSE}
statistics <- c("Minimum", "Maximum", "Mean")

#Temperature
Sensor_1 <- c(min(anomaly$Sensor1_T), max(anomaly$Sensor1_T), mean(anomaly$Sensor1_T))
Sensor_2 <- c(min(anomaly$Sensor2_T), max(anomaly$Sensor2_T), mean(anomaly$Sensor2_T))
Sensor_3 <- c(min(anomaly$Sensor3_T), max(anomaly$Sensor3_T), mean(anomaly$Sensor3_T))

#Convert to dataframe
stat_Temp <- data.frame(statistics, Sensor_1, Sensor_2,Sensor_3)

#Display table
knitr::kable(stat_Temp, caption = "Temperaturstatistik")
```

```{r, echo=FALSE}
#Relative air humidity
Zenzor_1 <- c(min(anomaly$Sensor1_LF), max(anomaly$Sensor1_LF), mean(anomaly$Sensor1_LF))
Zenzor_2 <- c(min(anomaly$Sensor2_LF), max(anomaly$Sensor2_LF), mean(anomaly$Sensor2_LF))
Zenzor_3 <- c(min(anomaly$Sensor3_LF), max(anomaly$Sensor3_LF), mean(anomaly$Sensor3_LF))

#Convert to dataframe
stat_lfk <- data.frame(statistics, Zenzor_1, Zenzor_2, Zenzor_3)

#Display table
knitr::kable(stat_lfk, caption = "Luftfeuchtigkeitsstatistik", 
             col.names = c("statistics", "Sensor_1", "Sensor_2", "Sensor_3"))
```
