---
title: "Reference_Station_Zollikofen"
author: "Flueckiger_Rubens"
date: "2025-04-18"
output: 
  html_document:
    toc: true
---

In this Markdown characteristics of the reference station Bern Zollikofen are displayed

# 1 Importing packages, file and function

```{r, warning=FALSE, message=FALSE}
#Packages
library(ggplot2)
library(dplyr)
library(lubridate)
library(RColorBrewer)
library(scales)

#File
station <- read.table("../data/Messwerte.csv", sep = ";", header = T)

#Function
source("../Functions/wind_rose_diagram.R")
```

Source of Code (adjusted version):

Guilherme Araujo Lima da Silva, 2016. Wind rose with ggplot (R)?. Link: <https://stackoverflow.com/questions/17266780/wind-rose-with-ggplot-r>. Called: 21.04.2025. *Note*: Changes were done to rename Axis & Legend

# 2 The dataset

The dataset contains:

-   stn: Station name (3 letter)
-   date: Date of measuring, YYYYMMDD
-   time: Time of measuring, HH:MM
-   tre200s0: Air temperature 2m above ground [°C]
-   ure200s0: Relative air humidity 2m above ground [%]
-   sre200z0: Sunshine duration, 10min sum [min]
-   fu3010z0: Wind speed, mean over 10min [km/h]
-   dkl010z0: Wind direction, mean over 10min [°]

# 3 Dataset preparation

In this step following processes are performed:

-   remove stn
-   converting date into a date form
-   rename tre200s0 to temp
-   rename ure200s0 to lfk
-   rename sre000z0 to sundur
-   rename fu3010z0 to windspeed
-   rename dkl010z0 to winddir

```{r}
station <- select(station, -stn)
station <- mutate(station, date = ymd(date))
station <- rename(station, temp = tre200s0)
station <- rename(station, lfk = ure200s0) 
station <- rename(station, sundur = sre000z0)
station <- rename(station, windspeed = fu3010z0)
station <- rename(station, winddir = dkl010z0)
```

# 3 Create plots

In this step following plots are created:

-   Temperature evolution over time
-   Relative humidity evolution over time
-   Sunshine duration (sum) over 10 minutes
-   Windspeed (mean) over 10 min
-   winddirection (mean) over 10 min

Then save it as jpeg

```{r, echo=FALSE, warning=FALSE}
#Open jpeg
jpeg("../Images/Reference_Station_Zollikofen/Referenzstation.jpeg", height = 1000, width = 1200, res = 108) 

#Temperature
p_temp <- ggplot(data = station, aes(x = time, y = temp, group = F)) + #Add variables
  geom_line() + geom_point(shape = 16) + #Add points + line
  xlab("Zeit") + ylab("Temperatur [°C]") + ggtitle("Mittlere Temperatur über 10 Minuten") + #Add title + axis label
  annotate(geom = "label", x = 1, y = 0.98, label = "I", family = "serif", size = 6) + #Plot number
  theme_bw() + #Change Plot motive
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Relative Humidity
p_lfk <- ggplot(data = station, aes(x = time, y = lfk, group = F)) + #Add variables
  geom_line() + geom_point(shape = 16) + #Add points + line
  xlab("Zeit") + ylab("Luftfeuchtigkeit [%]") + ggtitle("Mittlere Luftfeuchtigkeit über 10 Minuten") + #Add title + axis label
  annotate(geom = "label", x = 13, y = 77.8, label = "II", family = "serif", size = 6) + #Plot number
  theme_bw() + #Change Plot motive
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Sun duration
p_sundur <- ggplot(data = station, aes(x = time, y = sundur, group = F)) + #Add variables
  geom_line() + geom_point(shape = 16) + #Add points + line
  xlab("Zeit") + ylab("Sonnenscheindauer [min]") + ggtitle("Sonnenscheindauer in 10 Minuten") + #Add title + axis label
  ylim(0, 10) +
  annotate(geom = "label", x = 1, y = 9.6, label = "III", family = "serif", size = 6) + #Plot number
  theme_bw() + #Change Plot motive
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Wind speed
p_windspeed <- ggplot(data = station, aes(x = time, y = windspeed, group = F)) + #Add variables
  geom_line() + geom_point(shape = 16) + #Add points + line
  xlab("Zeit") + ylab("Windgeschwindigkeit [km/h]") + ggtitle("Mittlere Windgeschwindigkeit über 10 Minuten") + #Add title + axis label
  annotate(geom = "label", x = 1, y = 4.15, label = "IV", family = "serif", size = 6) + #Plot number
  theme_bw() + #Change Plot motive
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Wind direction
p_winddir <- plot.windrose(data = station, spd = station$windspeed, dir = station$winddir, spdmin = 1, spdmax = 5, spdseq = c(1,2,3,4,5), palette = "RdBu") +
  ggtitle("Windrosendiagramm Zollikofen") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) + #Center title
  theme(axis.title.x.bottom = element_blank())

#Combine all plots
cowplot::plot_grid(p_temp, p_lfk, p_sundur, p_windspeed, p_winddir, nrow = 3, ncol = 2)

#Save jpeg
dev.off()
```

# 5 Statistics

Calculating mean, minimum and maximum value of each variable

```{r, echo=FALSE}
Statistics <- c("Minimum", "Maximum", "Mittelwert")

Temperatur <- c(min(station$temp), max(station$temp), mean(station$temp))
Rel_Luftfeuchtigkeit <- c(min(station$lfk), max(station$lfk), mean(station$lfk))
Sonnenscheindauer <- c(min(station$sundur), max(station$sundur), mean(station$sundur))
Windgeschwindigkeit <- c(min(station$windspeed), max(station$windspeed), mean(station$windspeed))

#Create dataframe
stat_zoll <- data.frame(Statistics, Temperatur, Rel_Luftfeuchtigkeit, Sonnenscheindauer, Windgeschwindigkeit)

#Create table
knitr::kable(stat_zoll, caption = "Statistik Messstation Zollikofen",
             col.names = c("Statistik", "Temperatur", "Rel. Luftfeuchtigkeit", "Sonnenscheindauer", "Windgeschwindikgeit"))
```
