---
title: "Statistics"
author: "Flueckiger_Rubens"
date: "2025-04-01"
output: 
  html_document:
    toc: true
---

This file plots statistics of all 4 sensors

# 1 Import packages and files

```{r, message=FALSE, warning=FALSE}
#Packages
library(dplyr)

#Files
s_lap <- read.table("../data/Distanz_&_Hohe_KR.csv", sep = ";", header = T)
l_lap <- read.table("../data/Distanz_&_Hohe_GR.csv", sep = ";", header = T)
```

# 2 Information about sensors

1.  0.5m above ground, colour = red
2.  1m above ground, colour = blue
3.  2m above ground, colour = green
4.  3m above ground, colour = orange

```{r, echo=FALSE}
Sensor <- c(1, 2, 3, 4)
Height_sensor_m <- c(0.5, 1, 2, 3)
Colour <- c("Rot","Blau","Grün","Orange")
Info_Sensors <- data.frame(Sensor, Height_sensor_m, Colour)

#Table
knitr::kable(Info_Sensors, capture = "Information über Messstationen",
             col.names = c("Sensor", "Höhe", "Farbe"))
```

# 3 Statistics about each sensor

Mean quadratic deviation from sensor_1

To solve this problem, first a new dataset *anomaly* gets created. It contains:

-   Messung: This is the number of measurment
-   Reference sensor: Sensor_3 as reference (all values are = 0)
-   Remaining sensors: (sensor - reference sensor)$^2$
-   Both reference sensor and remaining sensors for temperature as well as for relative air humidity

```{r, echo=FALSE}
#Creating data frame
Messung <- l_lap$Messung
anomaly <- data.frame(Messung)

#Adding Values for temperature
anomaly <- mutate(anomaly, Sensor1_T = (l_lap$Sensor1_T - l_lap$Sensor4_T)^2) %>% #Sensor_0.5
  mutate(anomaly, Sensor2_T = (l_lap$Sensor2_T - l_lap$Sensor4_T)^2) %>% #Sensor_1
  mutate(anomaly, Sensor3_T = (l_lap$Sensor3_T - l_lap$Sensor4_T)^2) %>% #Sensor_2
  mutate(anomaly, Sensor4_T = (l_lap$Sensor4_T - l_lap$Sensor4_T)^2) #Sensor_3

#Adding Values for relative air humidity
anomaly <- mutate(anomaly, Sensor1_LF = (l_lap$Sensor1_LF - l_lap$Sensor4_LF)^2) %>% #Sensor_0.5
  mutate(anomaly, Sensor2_LF = (l_lap$Sensor2_LF - l_lap$Sensor4_LF)^2) %>% #Sensor_1
  mutate(anomaly, Sensor3_LF = (l_lap$Sensor3_LF - l_lap$Sensor4_LF)^2) %>% #Sensor_2
  mutate(anomaly, Sensor4_LF = (l_lap$Sensor4_LF - l_lap$Sensor4_LF)^2) #Sensor_3
```

Calculating mean, minimum and maximum value of each sensor

Temperature

```{r, echo=FALSE}
Statistics <- c("Minimum", "Maximum", "Mean","Variance_to_Sensor_3")
Sensor_1t <- c(round(min(l_lap$Sensor1_T), 2),
               round(max(l_lap$Sensor1_T), 2),
               round(mean(l_lap$Sensor1_T), 2),
               round(mean(anomaly$Sensor1_T), 2))

Sensor_2t <- c(round(min(l_lap$Sensor2_T), 2),
               round(max(l_lap$Sensor2_T), 2),
               round(mean(l_lap$Sensor2_T), 2),
               round(mean(anomaly$Sensor2_T),2))

Sensor_3t <- c(round(min(l_lap$Sensor3_T), 2),
               round(max(l_lap$Sensor3_T), 2),
               round(mean(l_lap$Sensor3_T), 2),
               round(mean(anomaly$Sensor3_T), 2))

Sensor_4t <- c(round(min(l_lap$Sensor4_T), 2),
               round(max(l_lap$Sensor4_T), 2),
               round(mean(l_lap$Sensor4_T), 2),
               round(mean(anomaly$Sensor4_T), 2))

#Create dataframe
stat_T <- data.frame(Statistics, Sensor_1t, Sensor_2t, Sensor_3t, Sensor_4t)

#Create table
knitr::kable(stat_T, caption = "Statistik Temperatur",
             col.names = c("Statistik", "Sensor 1", "Sensor 2", "Sensor 3", "Sensor 4"))
```

Relative air humidity

```{r, echo=FALSE}
Sensor_1lfk <- c(round(min(l_lap$Sensor1_LF), 2),
                 round(max(l_lap$Sensor1_LF), 2),
                 round(mean(l_lap$Sensor1_LF), 2),
                 round(mean(anomaly$Sensor1_LF), 2))

Sensor_2lfk <- c(round(min(l_lap$Sensor2_LF), 2),
                 round(max(l_lap$Sensor2_LF), 2),
                 round(mean(l_lap$Sensor2_LF), 2),
                 round(mean(anomaly$Sensor2_LF), 2))

Sensor_3lfk <- c(round(min(l_lap$Sensor3_LF), 2),
                 round(max(l_lap$Sensor3_LF), 2),
                 round(mean(l_lap$Sensor3_LF), 2),
                 round(mean(anomaly$Sensor3_LF), 2))

Sensor_4lfk <- c(round(min(l_lap$Sensor4_LF), 2),
                 round(max(l_lap$Sensor4_LF), 2),
                 round(mean(l_lap$Sensor4_LF), 2),
                 round(mean(anomaly$Sensor4_LF), 2))

#Create table
stat_LFK <- data.frame(Statistics,Sensor_1lfk,Sensor_2lfk,Sensor_3lfk,Sensor_4lfk)

#Create table
knitr::kable(stat_LFK, caption = "Statistik Temperatur",
             col.names = c("Statistik", "Sensor 1", "Sensor 2", "Sensor 3", "Sensor 4"))
```
