---
title: "Vertical_Profile"
author: "Flueckiger_Rubens"
date: "2025-05-27"
output: 
  html_document:
    toc: true
---

In this file the vertical profiles of temperature and relative humidity gets calculated

*Note:* If a better visual performance of the graphic is needed, I recommend using the Julia script provided in the folder *Julia*. Dataset needed as well as the saved grafics are provided inside that folder. 

# 1 Import packages, functions and files

```{r, warning=FALSE, message=FALSE}
#Packages
library(ggplot2)
library(reshape2)

#Files
data <- read.table("../data/Distanz_&_Hohe_GR.csv", sep = ";", header = T)

#Functions
source("../Functions/cubic_parameters.R")
```

# 2 Colours

Set colour codes for the plots and the legend 

```{r, echo=FALSE}
colors = c("Spannweite" = "red", "Mittelwert" = "black", "Stadt" = "grey70", "Wald" = "darkgreen")
```

# 3 Create new dataset

To improve runtime and visual aesthetics the initial 3894 data points of each sensor along the long route is reduced to 354 data points. This gets accomplished by taking an average over 11 consecutive data points.

```{r, echo=FALSE}
Messung <- seq(1, 354) #Number of new observations

#New dataframes
temp <- data.frame(Messung)
lfk <- data.frame(Messung)

#Mean values over 11 observations
for(i in 1:354){
  #Local climate zone
  lcz <- mean(data[seq(11 * i - 10, i * 11), "LCZ"])
  if(lcz > 1.5){ #If the value is bigger than 1.5 it is most probably forest
    lcz <- 2
  }
  else{ #If not urban
    lcz <- 1
  }
  
  #Mean values of temperature
  temp[i, "Sensor05_T"] <- mean(data[seq(11 * i - 10, i * 11), "Sensor1_T"])
  temp[i, "Sensor1_T"] <- mean(data[seq(11 * i - 10, i * 11), "Sensor2_T"])
  temp[i, "Sensor2_T"] <- mean(data[seq(11 * i - 10, i * 11), "Sensor3_T"])
  temp[i, "Sensor3_T"] <- mean(data[seq(11 * i - 10, i * 11), "Sensor4_T"])
  temp[i, "LCZ"] <- lcz
  
  #Mean values of humidity
  lfk[i, "Sensor05_LF"] <- mean(data[seq(11 * i - 10, i * 11), "Sensor1_LF"])
  lfk[i, "Sensor1_LF"] <- mean(data[seq(11 * i - 10, i * 11), "Sensor2_LF"])
  lfk[i, "Sensor2_LF"] <- mean(data[seq(11 * i - 10, i * 11), "Sensor3_LF"])
  lfk[i, "Sensor3_LF"] <- mean(data[seq(11 * i - 10, i * 11), "Sensor4_LF"])
  lfk[i, "LCZ"] <- lcz
}
```

# 3 Prepare plot

## 3.1 Cubic parameters

In a first step the function cubic parameters is applied to the temperature and relative air humidity values. Aim of this step is to calculate the cubic parameters a, b, c and d of the general cubic function (see below).
$$f(x) = ax^3 + bx^2 + cx + d$$

```{r, echo=FALSE}
#Apply function to temperature values
vert_temp <- cubic_parameters(half_meter = temp$Sensor05_T,
                              meter = temp$Sensor1_T,
                              two_meter = temp$Sensor2_T,
                              three_meter = temp$Sensor3_T)
vert_temp["LCZ"] <- temp$LCZ

#Apply function to humidity values
vert_lfk <- cubic_parameters(half_meter = lfk$Sensor05_LF,
                             meter = lfk$Sensor1_LF,
                             two_meter = lfk$Sensor2_LF,
                             three_meter = lfk$Sensor3_LF)
vert_lfk["LCZ"] <- lfk$LCZ
```

Since extreme values over the whole sampling duration is concidered as well, the function *cubic_parameters* is applied over mean, maximum and minimum values for both temperature and humidity.

```{r, echo=FALSE}
#Max temperature
vert_t_max <- cubic_parameters(half_meter = max(temp$Sensor05_T),
                              meter = max(temp$Sensor1_T),
                              two_meter = max(temp$Sensor2_T),
                              three_meter = max(temp$Sensor3_T))

#Min temperature
vert_t_min <- cubic_parameters(half_meter = min(temp$Sensor05_T),
                              meter = min(temp$Sensor1_T),
                              two_meter = min(temp$Sensor2_T),
                              three_meter = min(temp$Sensor3_T))

#Mean temperature
vert_t_mean <- cubic_parameters(half_meter = mean(temp$Sensor05_T),
                              meter = mean(temp$Sensor1_T),
                              two_meter = mean(temp$Sensor2_T),
                              three_meter = mean(temp$Sensor3_T))

#Store each value into one dataframe
vert_t_extreme <- rbind(vert_t_max, vert_t_min, vert_t_max)

#Max humidity
vert_lfk_max <- cubic_parameters(half_meter = max(lfk$Sensor05_LF),
                              meter = max(lfk$Sensor1_LF),
                              two_meter = max(lfk$Sensor2_LF),
                              three_meter = max(lfk$Sensor3_LF))

#Min humidity
vert_lfk_min <- cubic_parameters(half_meter = min(lfk$Sensor05_LF),
                              meter = min(lfk$Sensor1_LF),
                              two_meter = min(lfk$Sensor2_LF),
                              three_meter = min(lfk$Sensor3_LF))

#Mean humidity
vert_lfk_mean <- cubic_parameters(half_meter = mean(lfk$Sensor05_LF),
                              meter = mean(lfk$Sensor1_LF),
                              two_meter = mean(lfk$Sensor2_LF),
                              three_meter = mean(lfk$Sensor3_LF))

#Store each value into one dataframe
vert_lfk_extreme <- rbind(vert_lfk_max, vert_lfk_min, vert_lfk_max)
```

## 3.2 Curves

After calculating the parameters the curves are created. This step involves creating points in a small interval such that these points can be connected by a line further.

In a first step the quote on quote "normal" points are converted into a curve then the extreme values of both temperature and humidity. The latter is stored inside a seperate dataframe agian such that it can be accessed more easily at the plotting step.

```{r, echo=FALSE}
#Temperature
h <- seq(0.5, 3, by = (3 - 0.5)/353)
plot_t <- data.frame(h) #Dataframe used for plotting

#Create curve of temperature
for(j in 1:354){
  plot_t[, j] <- vert_temp[j, "a"] * h^3 + 
  vert_temp[j, "b"] * h^2 + 
  vert_temp[j, "c"] * h + 
  vert_temp[j, "d"]
}

#Create dataframe for curve extreme values of temperature
plot_t_extreme <- data.frame(h)

#Create curve of max temperature
plot_t_extreme[, "max"] <- vert_t_max$a * h^3 +
  vert_t_max$b * h^2 +
  vert_t_max$c * h +
  vert_t_max$d

#Create curve of min temperature
plot_t_extreme[, "min"] <- vert_t_min$a * h^3 +
  vert_t_min$b * h^2 +
  vert_t_min$c * h +
  vert_t_min$d

#Create curve of mean temperature
plot_t_extreme[, "mean"] <- vert_t_mean$a * h^3 +
  vert_t_mean$b * h^2 +
  vert_t_mean$c * h +
  vert_t_mean$d
```

```{r, echo=FALSE}
#Humidity
h <- seq(0.5, 3, by = (3 - 0.5)/353)
plot_lfk <- data.frame(h) #Dataframe used for plotting

#Create curve of temperature
for(j in 1:354){
  plot_lfk[, j] <- vert_lfk[j, "a"] * h^3 + 
  vert_lfk[j, "b"] * h^2 + 
  vert_lfk[j, "c"] * h + 
  vert_lfk[j, "d"]
}

#Create dataframe for curve extreme values of temperature
plot_lfk_extreme <- data.frame(h)

#Create curve of max temperature
plot_lfk_extreme[, "max"] <- vert_lfk_max$a * h^3 +
  vert_lfk_max$b * h^2 +
  vert_lfk_max$c * h +
  vert_lfk_max$d

#Create curve of min temperature
plot_lfk_extreme[, "min"] <- vert_lfk_min$a * h^3 +
  vert_lfk_min$b * h^2 +
  vert_lfk_min$c * h +
  vert_lfk_min$d

#Create curve of mean temperature
plot_lfk_extreme[, "mean"] <- vert_lfk_mean$a * h^3 +
  vert_lfk_mean$b * h^2 +
  vert_lfk_mean$c * h +
  vert_lfk_mean$d
```

# 4 Plots

Finally the plots can be created. Hereby we split the process into three steps.

## 4.1 Normal points

In this step the normal points are plotted. With a for loop every data point is plotted into a plot. Inside the plot the program checks if the current data point is located in a urban or a forest environment. If the data point is located in a urban environment, the curve is colered grey, else green.

```{r, echo=FALSE}
#Plot temperature curves
p1 <- ggplot(data = plot_t)

#For urban LCZ
for(k in 2:354){
  if(temp[k, "LCZ"] == 1){
      #Sub dataset for plot
      data_2 <- h
      data_2 <- data.frame(h)
      
      data_2[, "obj"] <- plot_t[, k] #Select current column
      p1 <- p1 + geom_path(data = data_2, aes(x = obj, y = h, color = "Stadt")) #Plot column
  }
}

#For forest LCZ
for(k in 2:235){
    if(temp[k, "LCZ"] == 2){
      #Sub dataset for plot
      data_2 <- h
      data_2 <- data.frame(h)
      
      data_2[, "obj"] <- plot_t[, k] #Select current column
      p1 <- p1 + geom_path(data = data_2, aes(x = obj, y = h, color = "Wald")) #Plot column
  }
  
}
```

```{r, echo=FALSE}
#Plot humidity curves
p2 <- ggplot(data = plot_lfk)

#For urban LCZ
for(k in 2:354){
  if(lfk[k, "LCZ"] == 1){
      #Sub dataset for plot
      data_3 <- h
      data_3 <- data.frame(h)
      
      data_3[, "obj"] <- plot_lfk[, k] #Select current column
      p2 <- p2 + geom_path(data = data_3, aes(x = obj, y = h, color = "Stadt")) #Plot column
  }
}

#For forest LCZ
for(k in 2:235){
    if(lfk[k, "LCZ"] == 2){
      #Sub dataset for plot
      data_3 <- h
      data_3 <- data.frame(h)
      
      data_3[, "obj"] <- plot_lfk[, k] #Select current column
      p2 <- p2 + geom_path(data = data_3, aes(x = obj, y = h, color = "Wald")) #Plot column
  }
  
}
```

## 4.2 Extremes

In a second step, all extreme values are plotted. This includes mean (in black), minimum and maximum (both in red).

```{r, echo=FALSE}
#Temperature
p1 <- p1 +
  #Plot extreme values of temperature
  geom_path(data = plot_t_extreme, aes(x = min, y = h, color = "Spannweite"), size = 1.5) + #Min
  geom_path(data = plot_t_extreme, aes(x = max, y = h), color = "red", size = 1.5) + #Max
  geom_path(data = plot_t_extreme, aes(x = mean, y = h, color = "Mittelwert"), size = 1.5) #Mean

#Humidity
p2 <- p2 +
  #Plot extreme values of temperature
  geom_path(data = plot_lfk_extreme, aes(x = min, y = h, color = "Spannweite"), size = 1.5) + #Min
  geom_path(data = plot_lfk_extreme, aes(x = max, y = h), color = "red", size = 1.5) + #Max
  geom_path(data = plot_lfk_extreme, aes(x = mean, y = h, color = "Mittelwert"), size = 1.5) #Mean
```

## 4.3 Aesthetics and Save

In a final step, the plots are visually adjusted. This includes:

- Adding Axis label
- Adding title
- Adding Legend

Finally the plots are saved inside the *Images* folder

```{r, echo=FALSE}
#Make plot adjustment
p1 <- p1 +
  labs(y = "Höhe über Boden [m]", #Y-Axis label
       x = "Temperatur [°C]", #X-Axis label
       title = "Vertikales Temperaturprofil") + #Title
  scale_colour_manual(name = "Legende", values = colors) + #Legend
  theme_bw() + #Change plot motive
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Open jpeg
jpeg("../Images/Vert_height_profile/temp.jpeg", height = 486, width = 780, res = 108)

p1


#Close and save jpeg
dev.off()
```
```{r, echo=FALSE}
#Make plot adjustment
p2 <- p2 +
  #Plot extreme values of temperature
  geom_path(data = plot_lfk_extreme, aes(x = min, y = h, color = "Spannweite")) + #Min
  geom_path(data = plot_lfk_extreme, aes(x = max, y = h), color = "red") + #Max
  geom_path(data = plot_lfk_extreme, aes(x = mean, y = h, color = "Mittelwert")) + #Mean
  
  labs(y = "Höhe über Boden [m]", #Y-Axis label
       x = "Luftfeuchtigkeit [%]", #X-Axis label
       title = "Vertikales Profil der relativen Luftfeuchtigkeit") + #Title
  scale_colour_manual(name = "Legende", values = colors) + #Legend
  theme_bw() + #Change plot motive
  theme(plot.title = element_text(hjust = 0.5)) #Center title


#Open jpeg
jpeg("../Images/Vert_height_profile/humidity.jpeg", height = 486, width = 780, res = 108)

p2


#Close and save jpeg
dev.off()
```
```{r, echo=FALSE, fig.align='center', fig.cap="Fig. 1: Height profile of temperature (left) and humidity (right)", fig.width=10, fig.height=4}
cowplot::plot_grid(p1,p2)
```

# 5 Statistics

In this step, a, b, c and d values of both temperature and humidity are displayed.

```{r, echo=FALSE}
#Temperature
mean_t <- data.frame(1)
mean_t["Grösse"] <- c("Temperatur")
mean_t["a"] <- round(vert_t_mean$a, 2)
mean_t["b"] <- round(vert_t_mean$b, 2)
mean_t["c"] <- round(vert_t_mean$c, 2)
mean_t["d"] <- round(vert_t_mean$d, 2)

#Humidity
mean_lfk <- data.frame(1)
mean_lfk["Grösse"] <- c("Luftfeuchtigkeit")
mean_lfk["a"] <- round(vert_lfk_mean$a, 2)
mean_lfk["b"] <- round(vert_lfk_mean$b, 2)
mean_lfk["c"] <- round(vert_lfk_mean$c, 2)
mean_lfk["d"] <- round(vert_lfk_mean$d,2 )

#Combine and display
mean_stat <- rbind(mean_t, mean_lfk)
mean_stat <- dplyr::select(mean_stat, -X1)

knitr::kable(mean_stat, caption = "Kubische Parameter der mittleren Temperaturveränderung")
```

According to this table, following equation can be written for the mean temperature evolution:
$$\overline{T(h)} = `r mean_stat[1, "a"]`x^3 + `r mean_stat[1, "b"]`x^2 + `r mean_stat[1, "c"]`x + `r mean_stat[1, "d"]`$$

And for the relative air humidity:

$$\overline{\varphi(h)} = `r mean_stat[2, "a"]`x^3 + `r mean_stat[2, "b"]`x^2 + `r mean_stat[2, "c"]`x + `r mean_stat[2, "d"]`$$

